---
title: "p8105_hw1_af3341"
author: Alana Ferris
output: github_document
date: "2022-09-16"
---

# Problem 1

```{r load data}

data("penguins", package = "palmerpenguins")

```

```{r, echo = FALSE, message = FALSE}

library(tidyverse)

```

## Dataset Description:
The penguins dataset contains `r nrow(penguins)` observations of penguins from the species Adelie, Gentoo, and Chinstrap. Some of the `r ncol(penguins)` variables include the  islands the penguins are from (Biscoe, Dream, Torgersen), as well as different physiological measurements. For example, the flipper length was measured and the mean length was found to be `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE))` mm.

```{r scatterplot, echo = FALSE}

penguins %>%
  ggplot(aes(x = bill_length_mm,
             y = flipper_length_mm,
             color = species)) +
  geom_point() + 
  xlab("Bill Length (mm)") + ylab("Flipper Length (mm)") + 
  labs(color = "Species")

ggsave("hw1_penguinsscatterplot.jpeg")

```

# Problem 2

## New Data Frame Attempt
```{r data frame}

hw1_df = 
  hw1_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  var_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
    var_factor = factor(c("male", "female", "nonbinary", "male", "male", "nonbinary", "male", "male", "female", "nonbinary")),
  )

x <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")
vectorLength = length(x)
cat("Length of vector x is :", vectorLength)

y <- c("male", "female", "nonbinary", "male", "male", "nonbinary", "male", "male", "female", "nonbinary")
vectorLength = length(y)
cat("Length of vector y is :", vectorLength)

```

I created a data frame with:

* A random sample size of 10 from a standard normal distribution with `rnorm`
* A logical vector indicating whether elements of the sample are greater than 0 with `norm_samp_pos`
* A character vector of length 10 with `var_char`
* A factor vector of length 10, with 3 different factor “levels” with `var_factor`

## Taking the Mean of Each Variable

### Mean of norm_samp
```{r mean attempt of norm_samp}

mean(pull(hw1_df, norm_samp))

```

### Mean of norm_samp_pos
```{r mean attempt of norm_samp_pos}

mean(pull(hw1_df, norm_samp_pos))

```

### Mean of var_char
```{r mean attempt of var_char}

mean(pull(hw1_df, var_char))

```

### Mean of var_factor
```{r mean attempt of var_factor}

mean(pull(hw1_df, var_factor))

```

_Outcome_:
I was able to take the mean of variables `norm_samp`, but not of `var_char` or `var_factor` because they are not numeric or logical. Surprisingly, I was also able to take the mean of `norm_samp_pos`, a logical variable, even though it is not numeric.

## Applying as.numeric

```{r as.numeric, eval = FALSE}

var_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")
V <- factor(var_char)
as.numeric(V)

var_factor = factor(c("male", "female", "nonbinary", "male", "male", "nonbinary", "male", "male", "female", "nonbinary"))
as.numeric (var_factor)

norm_samp = rnorm(10)
norm_samp_pos = norm_samp > 0
as.numeric(norm_samp_pos)

```

_Outcome_: In order to convert the character variable into a numeric variable via `as.numeric`, I first had to convert the character variable into a factor variable. This is because character variables can have an infinite number of responses whereas factor variables are nominal, meaning the response options are predefined as a number of levels. Factor variables have a different number of "levels", AKA number of responses, depending on the dataset. These levels can then correspond to numeric outputs once the `as.numeric` function is applied. For example, the factor variable `var_factor` has three factor levels (male, female, nonbinary) which can correspond to the numeric values 1, 2, 3. Character variables thus need to be given a defined set of levels before they can be processed as numeric. 
This also helps explain why we were able to take the mean of a logical but non-numeric variable, like `norm_samp_pos`, because it is a binary variable; the only options are 0 or 1 (true or false). Thus, you can immediately take the mean of a logical variable because they have inherent numeric value. Whereas factor and character variables first have to be explicitly converted into numeric values before a mean can be taken because they have more than 2 levels. 


